---
title: "Introduce FARS package"
author: "Duy Tho Nguyen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the FARS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include = FALSE}
library(fars)
library(dplyr)
library(maps)
```

The package provide functions used for sumarizing data from [FARS - Fatality Analysis Reporting System](http://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS))

# Install package

## Prerequisite

The package required `dplyr` and `maps` installed beforehand

```{r, echo=TRUE, eval=FALSE}
install.packages("dplyr")
install.packages("maps")
```

## Install from github

The package is currently not in CRAN but you could download from Github

```{r, eval=FALSE}
devtools::install_github("ThoDuyNguyen/fars")
```

# Usage

## Load packages

```{r, eval=FALSE}
library(fars)
library(dplyr)
library(maps)
```

## Package functions

We will use data from FARS, years 2013-2015, that can be found in the package directory under `\extdata` folder in the example.

## Read FARS data

`fars_read` read data from a file path provided as an argument. It returns a `tibble` with loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2014.csv.bz2", package = "fars")
fars_read(filename)
```

## Summarize number

`fars_summarize_years` takes a vector or list of years (numeric/integer values) as an argument and returns a pivot table with number of accidents per year and month. 

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "fars"))
fars_summarize_years(2013:2014)
```

## Plot accident locations

`fars_map_state` takesstate index and year as arguments and returns a plot of the state with accident locations. The state indices need to correspond available indeces from the data set. 
```{r fars_map_state_example}
setwd(system.file("extdata", package = "fars"))
fars_map_state(45, 2014)
```

 
